Flexible Scientific Data Management for Plant Phenomics Research

### 2013-03-11 s4biodiv2013/PODD-S4BIODIV-Submission.tex
comment out reference to PlantScan database
more fine tuning

%Affiliation:
\institute{
      eResearch Group, School of Information Technology and Electronic
Engineering,
             University of Queensland, Brisbane, Australia
 \and CSIRO Plant industry, High Resolution Plant Phenomics Centre, Canberra,
Australia
 \and CSIRO IMT Advanced Scientific Computing and Research Data Services,
Melbourne, Australia}

@@

\section{Introduction}
Since the genomics era, biology has become a data-driven science. Advances in
robotics, automation and imaging, in combination with high performance computing
have permitted the rapid production of large and complex biological datasets.
Currently, high volumes of heterogeneous image data, physiological and
morphological measurements are being acquired by a range of new phenotyping
platforms located in purpose built phenomics centres across the world. These
large datasets of phenotypic characters such as growth rate, plant architecture,
photosynthetic performance, yield must be stored and related to genotype. This
link is essential for expediting the identification of genetic variation in
natural and derived genetic populations (e.g. germplasm collections, association
genetic panels, recombinant inbred lines), understanding the dynamic
relationship between phenotype, genotype and environment and delivering the
increase in productivity necessary for feeding the world.

To achieve this goal, the vast array of phenotypic data collected from a variety
of phenomics platforms must be combined with metadata explaining how the raw
information was collected. This process forms structured digital objects
containing the raw information. These objects are then delivered to a range of
analysis pipelines, which transform the raw information into  aggregated
multi-phase datasets, each phase representing a new aggregation or inference
from the original raw data. This reduction process converts the raw
multi-dimensional data into information which is conceptually interpretable by a
human being, i.e. new knowledge. The details about the steps taken are recorded
to give context to the data. 
% These links enable scientists to generate novel
% conclusions about the plants being phenotyped, including conclusions based on
% the genotype, plant growth environments, treatments, phenotyping platforms and
% the phenotyping and analysis processes.

@@

\subsection{Redesign of the Phenomics Ontology Driven Data repository}
The PODD repository relies on semantic web technologies to
manage phenomics data and metadata. Although both ontologies and mappings are
essential, in PODD it was necessary to build the system with a relaxed
ontological vocabulary to allow scientists to generate datasets before they were
certain about the semantic meaning of their work. In addition, it was necessary
for PODD to allow scientists to continue to maintain projects containing curated
scientific concepts alongside raw experimental data. The PODD repository was 
redesigned based on an evaluation of the original software \cite{Li2010} that 
found it was not able to scale sufficiently well to suit the HRPPC needs. The 
major differences to the software implemented by \cite{Li2010} is that projects 
are no longer the only supported top object type, and projects are not stored 
in multiple parts, as that approach was not able to scale as was originally hypothesised.

@@

\subsection{Metadata}
Each acquisition on PlantScan\textsuperscript{\texttrademark} includes metadata
(in addition to the raw data
streams), such as plant genus and species, project and experiment metadata, a
unique identifier for each image (Globally Unique Identifier), imaging angle,
environmental temperature of the imaging chamber, location of optical and colour
calibration datasets for each acquisition run, LiDAR (Light Detection and
Ranging Sensors) calibration files. The metadata associated with each
acquisition is automatically generated when setting up the configuration on the
platform. This information is paramount to validate and process the raw image
data, and for the post-processing phases. This information currently resides in
the PlantScan\textsuperscript{\texttrademark} database.

>>>>

%Affiliation:
\institute{
      eResearch Group, School of Information Technology and Electronic
Engineering,
             University of Queensland, Brisbane, Australia
 \and CSIRO Plant industry, High Resolution Plant Phenomics Centre, Canberra,
Australia
 \and CSIRO IM\&T Advanced Scientific Computing and Research Data Services,
Melbourne, Australia}

@@

\section{Introduction}
Since the genomics era, biology has become a data-driven science. Advances in
robotics, automation and imaging, in combination with high performance computing
have permitted the rapid production of large and complex biological datasets.
Currently, high volumes of heterogeneous image data, physiological and
morphological measurements are being acquired by a range of new phenotyping
platforms located in purpose built phenomics centres across the world. These
large datasets of phenotypic characteristics such as growth rate, plant architecture,
photosynthetic performance, yield must be stored and correlated with genotypes. These factors provide evidence of genetic variation in natural and derived genetic populations (e.g. germplasm collections, association genetic panels, recombinant inbred lines). They also enable a deeper understanding the dynamic relationship between phenotype, genotype and environment which is necessary to continue delivering the increase in productivity necessary for feeding the world.

The vast array of phenotypic data collected from a variety
of phenomics platforms must be combined with metadata explaining how the raw
data was collected. This combination of raw data and metadata are then delivered to a range of analysis pipelines, which transform the raw data into aggregated
multi-phase datasets, each phase representing a new aggregation or inference
from the original raw data. This reduction process converts the raw
multi-dimensional data into information which is conceptually interpretable by a
human being, i.e. new knowledge. The additional metadata describing the steps taken are recorded to give context to the data. 
% These links enable scientists to generate novel
% conclusions about the plants being phenotyped, including conclusions based on
% the genotype, plant growth environments, treatments, phenotyping platforms and
% the phenotyping and analysis processes.

@@

\subsection{Redesign of the Phenomics Ontology Driven Data repository}
The PODD repository relies on semantic web technologies to
manage phenomics data and metadata. Although both ontologies and mappings are
essential, in PODD it was necessary to build the system with a relaxed
ontological vocabulary to allow scientists to generate datasets before they were
certain about the semantic meaning of their work. In addition, it was necessary
for PODD to allow scientists to continue to maintain projects containing curated
scientific concepts alongside raw experimental data. The PODD repository was 
redesigned based on an evaluation of the original software \cite{Li2010} that 
found it was not able to scale sufficiently to suit the HRPPC needs. The 
major differences to the software implemented by \cite{Li2010} are that projects 
are no longer the only supported top object type, and projects are not stored 
in multiple parts, as that approach was not able to scale as was originally hypothesised.

@@

\subsection{Metadata}
Each acquisition on PlantScan\textsuperscript{\texttrademark} includes metadata
(in addition to the raw data
streams), such as plant genus and species, project and experiment metadata, a
unique identifier for each image (Globally Unique Identifier), imaging angle,
environmental temperature of the imaging chamber, location of optical and colour
calibration datasets for each acquisition run, LiDAR (Light Detection and
Ranging Sensors) calibration files. The metadata associated with each
acquisition is automatically generated when setting up the configuration on the
platform. This information is paramount to validate and process the raw image
data, and for the post-processing phases. 
% This information currently resides in
% the PlantScan\textsuperscript{\texttrademark} database.
